---
title: "Group Project"
author: "Thai Pham - T00727094"
date: "2024-11-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import necessary libraries

```{r install_packages, echo = FALSE, message = FALSE, warning = FALSE}
pkg_list = c("dplyr", "ggplot2", "tidyr", "readr", "fable", "feasts", "fpp3", "lubridate", 'fabletools',
             "purrr", "tibble", "stringr", "forcats", 'tsibble', "rugarch",
             "lubridate", "tsibble", "forecast", "caret", 'GGally',
             "randomForest", "e1071", "xgboost", "glmnet", "quantmod", "gridExtra", "lubridate", "tseries", "rsample", "prophet")

# Install and load required packages
for (pkg in pkg_list) {
  # Check if the package is installed
  if (!require(pkg, character.only = TRUE)) {
    # Install if not available
    install.packages(pkg, dependencies = TRUE)
    # Load the package after installation
    library(pkg, character.only = TRUE)
  }
}

```

# Loading data
```{r}
# Step 1: Load the data
data <- read.csv("CANSIM.csv")

# Step 2: Create tsibble data
Cansim <- data %>%
  mutate(
    Year = as.integer(substr(REF_DATE, 1, 4))  # Extract the first year as integer
  ) %>%
  select(
    Year,                      # Newly created Year column
    Geo = GEO,                 # Rename GEO column to Geo
    Components = Components.of.population.growth,  # Rename Components.of.population.growth to Components
    Value = VALUE              # Rename VALUE to Value
  )

Cansim <- as_tsibble(Can, index = Year, key = c(Geo, Components))

# View the data
Cansim
head(Cansim)

# Extract the data of interest
Can_Immigrant <- Cansim %>% filter (Geo == "Canada", Components == "Immigrants")

Can_Immigrant$Adjust_Value <- Can_Immigrant$Value / 10^5
tail(Can_Immigrant)
``` 

# 1. EDA
```{r}
# Check for missing values across all columns
colSums(is.na(Can_Immigrant))

# Plot the data
Can_Immigrant %>% autoplot(Value)
```
## Box-cox transformation
```{r}
# Apply Box-Cox transformation using Guerrero method
lambda <- BoxCox.lambda(Can_Immigrant$Value, method = "guerrero")  # Find optimal lambda using Guerrero
Can_Immigrant <- Can_Immigrant %>% mutate(boxcox_Value = BoxCox(Value, lambda))
lambda
tail(Can_Immigrant)

# Plot original vs Box-Cox transformed series
autoplot(Can_Immigrant, Value) + 
  ggtitle("Original immigrants") +
  labs(y = "People")

autoplot(Can_Immigrant, boxcox_Value) + 
  ggtitle("Box-Cox Transformed Immigrants (Guerrero)") +
  labs(y = "Box-Cox Transformed Immigrants")

# ACF Plot
Can_Immigrant %>% ACF(Value) %>%
  autoplot() + labs(subtitle = "ACF plot of original data")

# PACF Plot
Can_Immigrant %>% PACF(Value) %>%
  autoplot() + labs(subtitle = "PACF plot of original data")
```

## Stationary and differencing check for original data
```{r}
# Check the stationarity of TS using KPSS test
Can_Immigrant |>
  features(Value, unitroot_kpss)
# Way 2 of using KPSS
KPSSTest = kpss.test(Can_Immigrant$Value)
KPSSTest

# Check the stationarity of TS using Augmented Dickey-Fuller (ADF) test
adf_test <- adf.test(Can_Immigrant$Value, alternative = "stationary")
print(adf_test)

# Determine whether differencing is required
Can_Immigrant |>
  features(Value, unitroot_ndiffs)

# Determine whether seasonal differencing is required
Can_Immigrant |>
  features(Value, unitroot_nsdiffs)
```
*KPSS test*: the p-value is shown as 0.01 (and therefore it may be smaller than that), indicating that the null hypothesis is rejected. That is, the data are not stationary. 
*ADF test*: the p-value = 0.4518 (larger than 0.05), indicating that the null hypothesis is NOT rejected. That is, the data are not stationary. 

## Stationary and differencing check for box-cox data
```{r}
# Check the stationarity of TS using KPSS test
Can_Immigrant |>
  features(boxcox_Value, unitroot_kpss)
# Way 2 of using KPSS
KPSSTest = kpss.test(Can_Immigrant$Value)
KPSSTest

# Check the stationarity of TS using Augmented Dickey-Fuller (ADF) test
adf_test <- adf.test(Can_Immigrant$boxcox_Value, alternative = "stationary")
print(adf_test)

# Determine whether differencing is required
Can_Immigrant |>
  features(boxcox_Value, unitroot_ndiffs)

# Determine whether seasonal differencing is required
Can_Immigrant |>
  features(boxcox_Value, unitroot_nsdiffs)
```
## Do the first differencing on box-cox data and check again
```{r}
# Do the first differencing and check again
Can_Immigrant <- Can_Immigrant %>%
  mutate(diff_Value = difference(boxcox_Value)) 

Can_Immigrant|>
  features(diff_Value, unitroot_kpss)

Can_Immigrant

```
This time, the p-value is reported as 0.1 (and so it could be larger than that). We can conclude that the differenced data appear stationary.

## Plot the original data with transformed and differencing data
```{r}
Can_Immigrant |>
  transmute(
    `Value` = Value,  
    `adjusted_Value` = boxcox_Value,
    `Annual change in adjusted Value` = diff_Value
  ) |>
  pivot_longer(-Year, names_to = "Type", values_to = "Value") |>
  mutate(
    Type = factor(Type, levels = c(
      "Value",
      "adjusted_Value",
      "Annual change in adjusted Value"
    ))
  ) |>
  as_tsibble(index = Year) |>
  ggplot(aes(x = Year, y = Value)) +
  geom_line() +
  facet_grid(vars(Type), scales = "free_y") +
  labs(title = "Canadian Statistic of Immigrants Analysis", y = NULL)

Can_Immigrant

```
# 2. Fit benchmark models
```{r}
# Generate train and test data
Can_Immigrant_train <- Can_Immigrant %>% filter(Year <= 2018)
Can_Immigrant_test <- Can_Immigrant %>% filter(Year > 2018)

tail(Can_Immigrant_train)
Can_Immigrant_test

# Plot train and test data
Can_Immigrant_train %>%
  autoplot(diff_Value) +
  autolayer(Can_Immigrant_test, diff_Value, colour = "red") +
  labs(title = "Train and Test Data", y = "adjusted_Value")

# Fit model
Can_Immigrant_fit <- Can_Immigrant_train %>%
   model(
    Naive = NAIVE(diff_Value),                   #Naive model
    Drift = RW(diff_Value ~ drift()),            #Drift model
    Mean = MEAN(diff_Value),                     #Mean model
    RandomWalk= RW(diff_Value)                   #Random walk model              
  )

# Information of fit model
augment(Can_Immigrant_fit) 

# Accuracy of fit model 
accuracy(Can_Immigrant_fit)
```
## Residuals diagnostics
```{r warning = FALSE}
# Check residuals for Naive model
gg_tsresiduals(Can_Immigrant_fit %>% select(Naive))

# For Drift model
gg_tsresiduals(Can_Immigrant_fit %>% select(Drift))

# For Mean model
gg_tsresiduals(Can_Immigrant_fit %>% select(Mean)) 

# For Random Walk model
gg_tsresiduals(Can_Immigrant_fit %>% select(RandomWalk))
```
## Portmanteau test
```{r}
augment(Can_Immigrant_fit) %>% features(.resid, ljung_box, lag=10, dof=0)
```
None of the benchmark models pass the Ljung-box test!

## Forecast and evaluate accuracy
```{r warning = FALSE}
# Generate Forecast
Can_Immigrant_fc <- Can_Immigrant_fit %>%
  forecast(h = "5 years")

# Check for missing forecast values
any(is.na(Can_Immigrant_fc$.mean))

# Plot the forecast 
Can_Immigrant_fc %>%
  autoplot(Can_Immigrant, level = NULL) +
  guides(colour = guide_legend(title = "Forecast"))

# Plot the forecasts with prediction interval
autoplot(Can_Immigrant, diff_Value) +
  autolayer(Can_Immigrant_fc)

# Accuracy of each forecast model
accuracy(Can_Immigrant_fc, Can_Immigrant)

```
# 3. Fit other TS models
## ARIMA

```{r}
# Fit ARIMA model 
Arima_model <- Can_Immigrant_train |>
  model(
    Arima = ARIMA(Value)
  )

# Check the model summary
Arima_model |>
  glance()

report(Arima_model)

# Forecast the ARIMA model
Arima_model_forecast <- Arima_model |>
  forecast(h = 5)  

# Plot the forecast
Arima_model_forecast |>
  autoplot(Can_Immigrant) +
  labs(title = "ARIMA Forecast", y = "Value")
```

## NNAR
```{r cache = TRUE}
# Fit neural network autoregression (NNAR) model
NNAR_model <- Can_Immigrant_train %>%
  model(NNETAR(Value))

NNAR_model %>% glance()

report(NNAR_model)

# Forecast the ARIMA model
NNAR_model_forecast <- NNAR_model |>
  forecast(h = 5)  

# Plot the forecast
NNAR_model_forecast |>
  autoplot(Can_Immigrant) +
  labs(title = "NNAR Forecast", y = "Value")
```
## TSLM
```{r}
TSLM_model <- Can_Immigrant_train %>%
  model(tslm = TSLM(Value ~ trend()))

coef(TSLM_model)

# Forecasting the next 5 years 
forecast_values <- TSLM_model %>% forecast(h = 5)

# Plot the forecast
forecast_values |>
  autoplot(Can_Immigrant) +
  labs(title = "TSLM Forecast", y = "Value")

```
```{r}
# Create a dummy variable for the TR to PR event in 2021
# Step 2: Create the dummy variable for TR2PR event in 2021 (for the training data)
Can_Immigrant_train <- Can_Immigrant_train %>%
  mutate(Event_TR2PR = ifelse(Year == 2021, 1, 0)) %>%
  as_tsibble(index = Year)

# Step 3: Build the TSLM model with trend and Event_TR2PR components using only the training data
TSLM_model_train <- Can_Immigrant_train %>%
  model(tslm_new = TSLM(Value ~ trend() + Event_TR2PR))

# Step 4: Forecast the next 5 years (2024-2028), assuming Event_TR2PR is 0 after 2021
future_years <- tibble(Year = 2024:2028) %>%
  mutate(Event_TR2PR = 0)  # Set Event_TR2PR to 0 for the forecast period

# Combine the training data with the future years for forecasting
Can_Immigrant_extended <- bind_rows(Can_Immigrant_train, future_years)

# Step 5: Perform the forecasting
forecast_train <- TSLM_model_train %>% forecast(h = 5)

# Step 6: Plot the forecast
forecast_train %>%
  autoplot(Can_Immigrant_extended) +
  labs(title = "TSLM Forecast with TR2PR Event (Training Data up to 2018)", 
       y = "Value", x = "Year")
```






## Prophet
```{r}
# Convert data to Prophet's required format
prophet_data <- Can_Immigrant %>%
  select(Year, Value) %>%
  rename(ds = Year, y = Value)
```

```{r}
# Initialize and fit the Prophet model
prophet_model <- prophet(prophet_data)
```
```{r cache = TRUE}
# Create a future dataframe for 10 more years 
future <- make_future_dataframe(prophet_model, periods = 10, freq = "year")

# Predict the future values
forecast <- predict(prophet_model, future)
```

```{r}
# Plot the forecast
plot(prophet_model, forecast) +
  labs(title = "Prophet Forecast", x = "Year", y = "Value")

# Plot the components to understand trend and seasonality
prophet_plot_components(prophet_model, forecast)
```

```{r}
# Extract forecasted values for the test period
test_forecast <- forecast %>%
  filter(ds %in% Can_Immigrant_test$Year)

# Join with the test set to evaluate accuracy
results <- Can_Immigrant_test %>%
  rename(actual = Value) %>%
  left_join(test_forecast %>% select(ds, yhat), by = c("Year" = "ds"))

# Calculate accuracy metrics (e.g., MAE, RMSE)
library(Metrics)
mae(results$actual, results$yhat)
rmse(results$actual, results$yhat)
```


# 4. Evaluate models
```{r}
# Accuracy of the ARIMA model
accuracy_arima <- accuracy(arima_forecast)

# Accuracy of the ETS model
accuracy_ets <- accuracy(ets_forecast)

# Accuracy of the Random Walk model
accuracy_rw <- accuracy(rw_forecast)

# Combine all accuracy results into a table
accuracy_results <- bind_rows(
  accuracy_arima %>% mutate(Model = "ARIMA"),
  accuracy_ets %>% mutate(Model = "ETS"),
  accuracy_rw %>% mutate(Model = "Random Walk")
)

# Display the accuracy results
accuracy_results

```




