---
title: "Project"
author: "Thai Pham - T00727094"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading necessary libraries
```{r results='hide'}
library(fpp3)
library(ggcorrplot)
library(conflicted)
library(tseries)
library(forecast)
library(prophet)

conflicts_prefer(fpp3::accuracy)
conflicts_prefer(dplyr::filter)
conflicts_prefer(vars::VAR)
conflicts_prefer(dplyr::select)

```

# Loading data
```{r}
data <- read.csv("CANSIM.csv")
```

# 1. Prepare and plot data
```{r}
CAN <- data %>%
  mutate(
    Year = as.integer(substr(REF_DATE, 1, 4))
    ) %>%
  select(Year, Components.of.population.growth, VALUE) %>%
  pivot_wider(names_from = Components.of.population.growth, values_from = VALUE) 

# Rename the variable
CAN <- CAN %>%
  rename(ReEmigrants = "Returning emigrants")

# Extract desired columns
CAN_df <- CAN %>% 
  select(Year, Births, Deaths, Immigrants, Emigrants, ReEmigrants)

# Create tsibble data
CAN_ts <- CAN_df %>%
  as_tsibble(index = Year)

# Reshape data for plotting
CAN_ts_long <- CAN_ts %>% 
  gather(key = "Variable", value = "Value", -Year)

# Plot the data
ggplot(CAN_ts_long, aes(x = Year, y = Value, color = Variable)) + 
  geom_line() +
  labs(title = "Time Series of Indicators", 
       x = "Year", 
       y = "Value") +
  theme_minimal() +
  theme(legend.title = element_blank())

```
```{r}
# Reshape the data to a long format for faceting
CAN_long <- CAN_df %>%
  pivot_longer(cols = c(Births, Deaths, Emigrants, ReEmigrants), 
               names_to = "Variable", 
               values_to = "Value")

# Plot Immigrants against other variables
ggplot(CAN_long, aes(x = Immigrants, y = Value, color = Variable)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + 
  facet_wrap(~ Variable, scales = "free_y") + 
  labs(
    title = "Immigrants vs Other Population Indicators",
    x = "Number of Immigrants",
    y = "Value",
    color = "Variable"
  ) +
  theme_minimal()
```

# 2. Statistic analysis
## Correlation
```{r}
# Correlation matrix
cor_matrix <- CAN_df %>%
  select(-Year)%>%
   cor(use = "complete.obs")

cor_matrix

# Heatmap plotting
ggcorrplot(cor_matrix, lab = TRUE)
```
## Explore trend and seasonal pattern
```{r}
immigrant_ts <- ts(CAN_ts$Immigrants, start = 1972, frequency = 1) 
trend <- stats::filter(immigrant_ts, filter = rep(1/3, 3), sides = 2)
plot(immigrant_ts, main = "Data and trend", col = "blue")
lines(trend, col = "red", lwd = 2)
legend("topright", legend = c("Data", "Trend"), col = c("blue", "red"), lty = 1)

```
## ACF and PACF plot
```{r warning = FALSE}
# Plot ACF and PACF for original data
par(mfrow = c(2, 1))
ggAcf(CAN_ts$Immigrants, main = "ACF of Immigrants in original data")
ggPacf(CAN_ts$Immigrants, main = "PACF of Immigrants in original data")

# Plot ACF and PACF for transformed data
par(mfrow = c(2, 1)) # Set plot layout to 2 rows, 1 column
ggAcf(CAN_transform$Immigrants, main = "ACF of Immigrants in transformed data")
ggPacf(CAN_transform$Immigrants, main = "PACF of Immigrants in transformed data")

```
# 3. Box-cox transformation and differencing
## (a) ADF test to check stationary
```{r}
adf.test(CAN_ts$Immigrants, alternative = "stationary")
adf.test(CAN_ts$Births, alternative = "stationary")
adf.test(CAN_ts$Deaths, alternative = "stationary")
adf.test(CAN_ts$Emigrants, alternative = "stationary")
adf.test(CAN_ts$ReEmigrants, alternative = "stationary")
```

### a. Immigrants
```{r}
# Apply Box-Cox transformation using Guerrero method
lambda_Immigrant <- BoxCox.lambda(CAN_ts$Immigrants, method = "guerrero")  
CAN_ts <- CAN_ts %>% mutate(boxcox_Immigrants = BoxCox(Immigrants, lambda_Immigrant))

# Determine whether differencing is required
CAN_ts |>
  features(boxcox_Immigrants, unitroot_ndiffs)

# Determine whether seasonal differencing is required
CAN_ts |>
  features(boxcox_Immigrants, unitroot_nsdiffs)

# Do the first differencing
CAN_ts <- CAN_ts %>%  mutate(diff_boxcox_Immigrants = c(NA, diff(boxcox_Immigrants, lag = 1)))

# ADF test on transformed data
CAN_ts_no_na <- CAN_ts %>%
  filter(!is.na(diff_boxcox_Immigrants))

adf.test(CAN_ts_no_na$diff_boxcox_Immigrants, alternative = "stationary")

# KPSS test on transformed data
CAN_ts |>
  features(diff_boxcox_Immigrants, unitroot_kpss)
```
=> For Immigrants variable: do the first differencing on box-cox value.

### b. Births
```{r}
# Apply Box-Cox transformation using Guerrero method
lambda_Births <- BoxCox.lambda(CAN_ts$Births, method = "guerrero")  
CAN_ts <- CAN_ts %>% mutate(boxcox_Births = BoxCox(Births, lambda_Births))

# Determine whether differencing is required
CAN_ts |>
  features(boxcox_Births, unitroot_ndiffs)

# Determine whether seasonal differencing is required
CAN_ts |>
  features(boxcox_Births, unitroot_nsdiffs)

# ADF test on transformed data
adf.test(CAN_ts$boxcox_Births, alternative = "stationary")

# KPSS test on transformed data
CAN_ts |>
  features(boxcox_Births, unitroot_kpss)
```

=> For Births variable: Only do the box-cox transform

### c. Deaths
```{r}
# Apply Box-Cox transformation using Guerrero method
lambda_Deaths <- BoxCox.lambda(CAN_ts$Deaths, method = "guerrero")
CAN_ts <- CAN_ts %>% mutate(boxcox_Deaths = BoxCox(Deaths, lambda_Deaths))

# Determine whether differencing is required
CAN_ts |>
  features(boxcox_Deaths, unitroot_ndiffs)

# Determine whether seasonal differencing is required
CAN_ts |>
  features(boxcox_Deaths, unitroot_nsdiffs)

# Do the first differencing
CAN_ts <- CAN_ts %>%  mutate(diff_boxcox_Deaths = c(NA, diff(boxcox_Deaths, lag = 1)))

# ADF test on transformed data
CAN_ts_no_na <- CAN_ts %>%
  filter(!is.na(diff_boxcox_Deaths))

adf.test(CAN_ts_no_na$diff_boxcox_Deaths, alternative = "stationary")

# KPSS test on transformed data
CAN_ts |>
  features(diff_boxcox_Deaths, unitroot_kpss)
```
=> For Deaths variable: Do the first differencing on box-cox value.

### d. Emigrants
```{r}
# Apply Box-Cox transformation using Guerrero method
lambda_Emigrants <- BoxCox.lambda(CAN_ts$Emigrants, method = "guerrero")  
CAN_ts <- CAN_ts %>% mutate(boxcox_Emigrants = BoxCox(Emigrants, lambda_Emigrants))

# Determine whether differencing is required
CAN_ts |>
  features(boxcox_Emigrants, unitroot_ndiffs)

# Determine whether seasonal differencing is required
CAN_ts |>
  features(boxcox_Emigrants, unitroot_nsdiffs)

# Do the first differencing
CAN_ts <- CAN_ts %>%  mutate(diff_boxcox_Emigrants = c(NA, diff(boxcox_Emigrants, lag = 1)))

# ADF test on transformed data
CAN_ts_no_na <- CAN_ts %>%
  filter(!is.na(diff_boxcox_Emigrants))

adf.test(CAN_ts_no_na$diff_boxcox_Emigrants, alternative = "stationary")

# KPSS test on transformed data
CAN_ts |>
  features(diff_boxcox_Emigrants, unitroot_kpss)
```
=> For Emigrants variable: Do the first differencing on box-cox value

### e. Returning Emigrants
```{r}
# Apply Box-Cox transformation using Guerrero method
lambda_ReEmigrant <- BoxCox.lambda(CAN_ts$ReEmigrants, method = "guerrero") 
CAN_ts <- CAN_ts %>% mutate(boxcox_ReEmigrant = BoxCox(ReEmigrants, lambda_ReEmigrant))

# Determine whether differencing is required
CAN_ts |>
  features(boxcox_ReEmigrant, unitroot_ndiffs)

# Determine whether seasonal differencing is required
CAN_ts |>
  features(boxcox_ReEmigrant, unitroot_nsdiffs)

# Do the first differencing
CAN_ts <- CAN_ts %>%  mutate(diff_boxcox_ReEmigrant = c(NA, diff(boxcox_ReEmigrant, lag = 1)))

# ADF test on transformed data
CAN_ts_no_na <- CAN_ts %>%
  filter(!is.na(diff_boxcox_ReEmigrant))

adf.test(CAN_ts_no_na$diff_boxcox_ReEmigrant, alternative = "stationary")

# KPSS test on transformed data
CAN_ts |>
  features(diff_boxcox_ReEmigrant, unitroot_kpss)
```
=> For Returning Emigrants variable: Do the first differencing on box-cox value

## (b) Transform data for analysis
```{r}
CAN_transform <- CAN_ts %>%
  select(Year, boxcox_Births, diff_boxcox_Immigrants, diff_boxcox_Deaths, diff_boxcox_Emigrants, diff_boxcox_ReEmigrant) %>%
  rename(
    Births = boxcox_Births,
    Immigrants = diff_boxcox_Immigrants,
    Deaths = diff_boxcox_Deaths,
    Emigrants = diff_boxcox_Emigrants,
    ReEmigrants = diff_boxcox_ReEmigrant
  ) %>%
  filter (Year > 1971) %>%
  as_tsibble(index = Year)
```


# 4. Fit benchmark models (on transformed data)
## Split data into train and test set
```{r}
# Generate train and test data
CAN_transform_train <- CAN_transform %>% filter(Year <= 2018)
CAN_transform_test <- CAN_transform %>% filter(Year > 2018)

# Plot train and test data
CAN_transform_train  %>%
  autoplot(Immigrants) +
  autolayer(CAN_transform_test, Immigrants, colour = "red") +
  labs(title = "Train and Test Data", y = "adjusted_Immigrants")
```

## Fit models on train and test set
```{r}
# Fit model on train set
models <- CAN_transform_train %>%
  model(
    Mean = MEAN(Immigrants),
    Naive = NAIVE(Immigrants),
    RandomWalk = RW(Immigrants),
    Drift = RW(Immigrants ~ drift())
  )

# Forecast on test set
forecasts <- models %>%
  forecast(h = nrow(CAN_transform_test))  

# Display the results
forecasts %>%
  autoplot(CAN_transform_train, level = NULL) +
  autolayer(CAN_transform_test, Immigrants, colour = "blue") +
  labs(title = "Forecasts vs Actual Data", y = "Immigrants")

# Accuracy on test set
accuracy(forecasts, CAN_transform_test)
```

## Fit models using CV
```{r}
set.seed(123)
CAN_transform_cv <- CAN_transform |>
  stretch_tsibble(.init = 30, .step = 1) |>
  filter(.id != max(.id))

# Evaluate models using CV accuracy
benchmark_cv <- CAN_transform_cv %>%
  model(
    Mean = MEAN(Immigrants),
    Naive = NAIVE(Immigrants),
    RandomWalk = RW(Immigrants),
    Drift = RW(Immigrants ~ drift())
  )|>
  forecast(h = 1)

accuracy(benchmark_cv, CAN_transform_test)

```

# 5. Fit other TS models (without Dynamic)
## (a) ARIMA (normal)
```{r warning = FALSE}
# Fit ARIMA model 
arima_fit <- CAN_transform_train %>%
   model(Arima = ARIMA(Immigrants)
  )

report(arima_fit)

# Check residuals for ARIMA model
gg_tsresiduals(arima_fit)

# Portmanteau test
augment(arima_fit) %>% features(.resid, ljung_box, lag=10, dof=0)

# Time series cross-validation accuracy
CAN_transform_cv %>%
  model (Arima = ARIMA(Immigrants)) %>% 
  forecast(h = 1) %>%
  accuracy(CAN_transform_test)

# Produce forecasts
forecast_ARIMA <- arima_fit %>%
  forecast(new_data = CAN_transform_test)

# Evaluate forecast accuracy
ARIMA_accuracy <- forecast_ARIMA %>%
  accuracy(CAN_transform_test)

ARIMA_accuracy

# Plot the forecasts along with the original data
forecast_ARIMA  %>%
  autoplot(CAN_transform, level = NULL) +
  labs(
    title = "Forecasts for ARIMA Model",
    y = "Immigrants",
    x = "Year"
  ) +
  theme_minimal()
```
## (b) NNAR (normal)
```{r warning = FALSE cache = TRUE}
# Fit NNAR model 
nnar_fit <- CAN_transform_train %>%
   model(nnar = NNETAR(Immigrants)
  )

report(nnar_fit)

# Check residuals for NNAR model
gg_tsresiduals(nnar_fit)

# Portmanteau test
augment(nnar_fit) %>% features(.resid, ljung_box, lag=10, dof=0)

# Time series cross-validation accuracy
CAN_transform_cv %>%
  model (nnar = NNETAR(Immigrants)) %>% 
  forecast(h = 1) %>%
  accuracy(CAN_transform_test)

# Produce forecasts
forecast_NNAR<- nnar_fit %>%
  forecast(new_data = CAN_transform_test)

# Evaluate forecast accuracy
NNAR_accuracy <- forecast_NNAR %>%
  accuracy(CAN_transform_test)

NNAR_accuracy

# Plot the forecasts along with the original data
forecast_NNAR  %>%
  autoplot(CAN_transform, level = NULL) +
  labs(
    title = "Forecasts for NNAR Model",
    y = "Immigrants",
    x = "Year"
  ) +
  theme_minimal()
```
## (c) TSLM (normal)

```{r}
# Fit TSLM model 
tslm_fit <- CAN_transform_train %>%
   model(tslm = TSLM(Immigrants)
  )

report(tslm_fit)

# Check residuals for TSLM model
gg_tsresiduals(tslm_fit)

# Portmanteau test
augment(tslm_fit) %>% features(.resid, ljung_box, lag=10, dof=0)

# Time series cross-validation accuracy
CAN_transform_cv %>%
  model (tslm = TSLM(Immigrants)) %>% 
  forecast(h = 1) %>%
  accuracy(CAN_transform_test)

# Produce forecasts
forecast_TSLM<- tslm_fit %>%
  forecast(new_data = CAN_transform_test)

# Evaluate forecast accuracy
TSLM_accuracy <- forecast_TSLM %>%
  accuracy(CAN_transform_test)

TSLM_accuracy

# Plot the forecasts along with the original data
forecast_TSLM  %>%
  autoplot(CAN_transform, level = NULL) +
  labs(
    title = "Forecasts for TSLM Model",
    y = "Immigrants",
    x = "Year"
  ) +
  theme_minimal()
```
## (d) PROPHET
```{r cache = TRUE}
# Prepare the data for Prophet
prophet_data <- CAN_transform_train %>%
  select(Year, Immigrants, Deaths, Emigrants, ReEmigrants) %>%
  rename(ds = Year, y = Immigrants)  

# Create Prophet model with external regressors
prophet_model <- prophet(
  yearly.seasonality = TRUE, 
  daily.seasonality = FALSE,
  weekly.seasonality = FALSE
)

# Add external regressors to the model before fitting
prophet_model <- add_regressor(prophet_model, 'Deaths')
prophet_model <- add_regressor(prophet_model, 'Emigrants')
prophet_model <- add_regressor(prophet_model, 'ReEmigrants')

# Fit the model
prophet_model <- fit.prophet(prophet_model, prophet_data)
```

```{r}
# Prepare the test data for forecasting
future_data <- CAN_transform_test %>%
  select(Year, Deaths, Emigrants, ReEmigrants) %>%
  rename(ds = Year)  

# Forecast for the test data (future values)
forecast <- predict(prophet_model, future_data)

# View the forecast results
tail(forecast)

# Plot the forecast
plot(prophet_model, forecast)

# Show the model components (trend, seasonality, and the effect of external regressors)
prophet_components <- prophet_plot_components(prophet_model, forecast)
prophet_components
```
```{r cache = TRUE}
actuals <- CAN_transform_test$Immigrants

# Extract the predicted values
predictions <- forecast$yhat

# Calculate MAE
mae <- mean(abs(predictions - actuals))

# Calculate MSE
mse <- mean((predictions - actuals)^2)

# Calculate RMSE
rmse <- sqrt(mse)

# Calculate MAPE
mape <- mean(abs((actuals - predictions) / actuals)) * 100


# Print the accuracy metrics
cat("MAE:", mae, "\n")
cat("MSE:", mse, "\n")
cat("RMSE:", rmse, "\n")
cat("MAPE:", mape, "%", "\n")
```
## (e) ETS
```{r}
# Prepare the data for ETS
CAN_transform_train_ts <- ts(CAN_transform_train$Immigrants, start = c(min(CAN_transform_train$Year)), frequency = 1)

# Fit the ETS model
ets_fit <- ets(CAN_transform_train_ts)

# Check residuals for the ETS model
checkresiduals(ets_fit)

# Compute accuracy on the training data
accuracy_train <- accuracy(ets_fit)

# Produce forecasts for the future period 
forecast_ets <- forecast(ets_fit, h = length(CAN_transform_test$Immigrants))

# Evaluate forecast accuracy on the test set
ets_accuracy <- accuracy(forecast_ets, CAN_transform_test$Immigrants)

# Print forecast accuracy
print(ets_accuracy)

# Plot the forecasts along with the original data
autoplot(forecast_ets) +
  geom_line(data = CAN_transform_train, aes(x = Year, y = Immigrants), color = "blue", size = 1) +
  labs(
    title = "Forecasts for ETS Model",
    y = "adjusted Immigrants",
    x = "Year"
  ) +
  theme_minimal()
```
## (f) EWMA
```{r}
# Define EWMA function
EWMA_forecast <- function(data, alpha, forecast_horizon) {
  n <- length(data)
  smoothed_values <- numeric(n)
  smoothed_values[1] <- data[1] # Initialize with the first value
  
  # Calculate EWMA
  for (i in 2:n) {
    smoothed_values[i] <- alpha * data[i] + (1 - alpha) * smoothed_values[i - 1]
  }
  
  # Forecast future values using the last smoothed value
  future_values <- rep(smoothed_values[n], forecast_horizon)
  
  list(smoothed = smoothed_values, forecast = future_values)
}
```

```{r}
# Define a range of alpha values to test
alpha_values <- seq(0.01, 0.4, by = 0.01)

# Initialize an empty tibble to store results
results <- tibble(
  alpha = numeric(),
  MAE = numeric(),
  RMSE = numeric(),
  MAPE = numeric()
)

# Loop through alpha values to calculate accuracy
for (alpha in alpha_values) {
  # Apply EWMA with the current alpha
  EWMA_result <- EWMA_forecast(CAN_transform_train$Immigrants, alpha, length(CAN_transform_test$Immigrants))
  
  # Calculate accuracy for the test set
  actual_test <- CAN_transform_test$Immigrants
  forecast_test <- EWMA_result$forecast
  
  MAE <- mean(abs(actual_test - forecast_test), na.rm = TRUE)
  RMSE <- sqrt(mean((actual_test - forecast_test)^2, na.rm = TRUE))
  MAPE <- mean(abs((actual_test - forecast_test) / actual_test) * 100, na.rm = TRUE)
  
  # Add results for the current alpha to the tibble
  results <- results %>%
    add_row(alpha = alpha, MAE = MAE, RMSE = RMSE, MAPE = MAPE)
}

# Find the best alpha based on the minimum RMSE
best_alpha <- results %>% filter(RMSE == min(RMSE)) %>% pull(alpha)

# Print the results
print(results)

# Print the best alpha
cat("Best alpha based on RMSE:", best_alpha, "\n")

# Apply the best alpha for final forecast
EWMA_best <- EWMA_forecast(CAN_transform_train$Immigrants, best_alpha, length(CAN_transform_test$Immigrants))

# Combine original data and forecasts for plotting
EWMA_forecasted <- tibble(
  Year = c(CAN_transform_train$Year, CAN_transform_test$Year),
  Immigrants = c(CAN_transform_train$Immigrants, rep(NA, length(CAN_transform_test$Immigrants))),
  EWMA = c(EWMA_best$smoothed, EWMA_best$forecast)
)

# Plot the EWMA results with the best alpha
EWMA_forecasted %>%
  ggplot(aes(x = Year)) +
  geom_line(aes(y = Immigrants), color = "blue", size = 1.2, linetype = "dashed", na.rm = TRUE) +
  geom_line(aes(y = EWMA), color = "red", size = 1.2) +
  labs(
    title = paste("EWMA Forecasting (Best Alpha =", best_alpha, ")"),
    y = "Immigrants",
    x = "Year"
  ) +
  theme_minimal()
```

## (g) TSLM with knot 2021
```{r warning = FALSE}
# Fit TSLM model with knot at year 2021
TSLM_fit_knot <- CAN_transform_train %>%
  model(piecewise = TSLM(Immigrants ~ trend(knots = 2021))
  )
 
# Check residuals for TSLM model with knot
gg_tsresiduals(TSLM_fit_knot)

# Portmanteau test for TSLM model residuals
augment(TSLM_fit_knot) %>% features(.resid, ljung_box, lag = 10, dof = 0)

# Time series cross-validation accuracy for TSLM with knot
CAN_transform_cv %>%
  model(piecewise = TSLM(Immigrants ~ trend(knots = 2021))) %>%
  forecast(h = 5) %>%
  accuracy(CAN_transform_test)

# Produce forecasts with TSLM with knot
forecast_TSLM_knot <- TSLM_fit_knot %>%
  forecast(new_data = CAN_transform_test)

# Evaluate forecast accuracy for TSLM with knot
TSLM_accuracy_knot <- forecast_TSLM_knot %>%
  accuracy(CAN_transform_test)

TSLM_accuracy_knot

# Plot the forecasts along with the original data for TSLM with knot
forecast_TSLM_knot %>%
  autoplot(CAN_transform, level = NULL) +
  labs(
    title = "Forecasts for TSLM Model with Knot at 2021",
    y = "Value",
    x = "Year"
  ) +
  theme_minimal()
```

# 6. Fit Dynamic TS models (NOTE: data has already be transformed!!!)
## (a) ARIMA
```{r}
# Fit ARIMA with external regressors
arima_model <- CAN_transform_train %>%
  model(
    ARIMA_Regressors = ARIMA(Immigrants ~  Deaths + Emigrants + ReEmigrants)
  )

# Summary of the fitted model
report(arima_model)


# Forecasting for the test set period (2019-2023)
arima_forecast <- arima_model %>%
  forecast(new_data = CAN_transform_test)

# Plot the forecast
arima_forecast %>%
  autoplot(CAN_transform, level = NULL) +
  autolayer(CAN_transform_test, Immigrants, colour = "red") +
  labs(
    title = "ARIMA with Regressors Forecast",
    y = "Immigrants",
    x = "Year"
  ) +
  theme_minimal()

# Evaluate forecast accuracy
accuracy_arima <- accuracy(arima_forecast, CAN_transform_test)
accuracy_arima

# Residual diagnostics
arima_model %>%
  gg_tsresiduals()

# Ljung-Box test for residual autocorrelation
augment(arima_model) %>%
  features(.innov, ljung_box, lag = 10)

```
## (b) NNAR
```{r}
# Fit NNAR model with external regressors
nnar_model <- CAN_transform_train %>%
  model(
    NNAR_Regressors = NNETAR(Immigrants ~ Deaths + Emigrants + ReEmigrants, p=1)
  )

# Summary of the fitted NNAR model
report(nnar_model)
```
```{r cache = TRUE}
# Forecast on the test set with external regressors
nnar_forecasts <- nnar_model %>%
  forecast(new_data = CAN_transform_test)

# View the forecasted values
nnar_forecasts

# Plot the forecast
nnar_forecasts %>%
  autoplot(CAN_transform, level = NULL) +
  autolayer(CAN_transform_test, Immigrants, colour = "red") +
  labs(
    title = "NNAR with Regressors Forecast",
    y = "Immigrants",
    x = "Year"
  ) +
  theme_minimal()

# Accuracy of the forecast
nnar_accuracy <- accuracy(nnar_forecasts, CAN_transform_test)

# Print the accuracy results
print(nnar_accuracy)

# Residual diagnostics
nnar_model %>%
  gg_tsresiduals()

# Ljung-Box test for residual autocorrelation
augment(nnar_model) %>%
  features(.innov, ljung_box, lag = 10)
```
## (c) TSLM
```{r}
# Fit TSLM model with external regressors
tslm_model <- CAN_transform_train %>%
  model(
    TSLM_Regressors = TSLM(Immigrants ~ Deaths + Emigrants + ReEmigrants)
  )

# Summary of the fitted TSLM model
report(tslm_model)
```
```{r}
# Forecast on the test set with external regressors
tslm_forecasts <- tslm_model %>%
  forecast(new_data = CAN_transform_test)

# View the forecasted values
tslm_forecasts

# Plot the forecast
tslm_forecasts %>%
  autoplot(CAN_transform, level = NULL) +
  autolayer(CAN_transform_test, Immigrants, colour = "red") +
  labs(
    title = "TSLM with Regressors Forecast",
    y = "Immigrants",
    x = "Year"
  ) +
  theme_minimal()

# Accuracy of the forecast
tslm_accuracy <- accuracy(tslm_forecasts, CAN_transform_test)

# Print the accuracy results
print(tslm_accuracy)

# Residual diagnostics
tslm_model %>%
  gg_tsresiduals()

# Ljung-Box test for residual autocorrelation
augment(tslm_model) %>%
  features(.innov, ljung_box, lag = 10)
```

## (d) ARIMA with trend

```{r}
# Fit ARIMA with external regressors
arima_trend <- CAN_transform_train %>%
  model(
    ARIMA_Regressors = ARIMA(Immigrants ~  Deaths + Emigrants + ReEmigrants + trend())
  )

# Summary of the fitted model
report(arima_trend)
```
```{r}
# Forecasting for the test set period (2019-2023)
arima_trend_forecast <- arima_trend %>%
  forecast(new_data = CAN_transform_test)

# Plot the forecast
arima_trend_forecast %>%
  autoplot(CAN_transform, level = NULL) +
  autolayer(CAN_transform_test, Immigrants, colour = "red") +
  labs(
    title = "ARIMA with Regressors Forecast",
    y = "Immigrants",
    x = "Year"
  ) +
  theme_minimal()

# Evaluate forecast accuracy
accuracy_arima_trend <- accuracy(arima_trend_forecast, CAN_transform_test)
accuracy_arima_trend

# Residual diagnostics
arima_trend %>%
  gg_tsresiduals()

# Ljung-Box test for residual autocorrelation
augment(arima_trend) %>%
  features(.innov, ljung_box, lag = 10)

```

## (e) TSLM with knot + ARIMA errors (full TSLM manually)
```{r}
# Step 1: Fit TSLM model
tslm_fit <- CAN_transform_train %>%
  model(tslm = TSLM(Immigrants ~ Deaths + Births + ReEmigrants + trend(knots = 2021)))

# Step 2: Extract residuals and prepare tsibble
residuals_tslm <- augment(tslm_fit) %>%
  select(Year, residuals = .resid) %>% 
  as_tsibble(index = Year)

# Step 3: Fit ARIMA model on residuals
arima_residuals <- residuals_tslm %>%
  model(arima = ARIMA(residuals ~ pdq(1,0,1)))

arima_residuals

# Step 4: Forecast ARIMA residuals
forecast_arima_residuals <- arima_residuals %>%
  forecast(h = nrow(CAN_transform_test))

# Step 5: Combine TSLM predictions with ARIMA residual forecasts
forecast_TSLM <- tslm_fit %>%
  forecast(new_data = CAN_transform_test)

combined_forecast <- forecast_TSLM %>%
  mutate(.mean = .mean + forecast_arima_residuals$.mean)

# Step 6: Evaluate combined forecast accuracy
accuracy_combined <- combined_forecast %>%
  accuracy(CAN_transform_test)

# Print results
print(accuracy_combined)

# Step 7: Plot combined forecasts
combined_forecast %>%
  autoplot(CAN_transform, level = NULL) +
  labs(
    title = "Forecasts for TSLM + ARIMA Errors",
    y = "Immigrants",
    x = "Year"
  ) +
  theme_minimal()

```
## (f) TSLM with knot at 2021

```{r warning = FALSE}
# Fit TSLM model with knot at year 2021
TSLM_dynamic__fit_knot <- CAN_transform_train %>%
  model(piecewise = TSLM(Immigrants ~ Deaths + Emigrants + ReEmigrants + trend(knots = 2021))
  )
 
# Check residuals for TSLM model with knot
gg_tsresiduals(TSLM_dynamic__fit_knot )

# Portmanteau test for TSLM model residuals
augment(TSLM_dynamic__fit_knot ) %>% features(.resid, ljung_box, lag = 10, dof = 0)


# Produce forecasts with TSLM with knot
forecast_TSLM_dynamic_knot <- TSLM_dynamic__fit_knot %>%
  forecast(new_data = CAN_transform_test)

# Evaluate forecast accuracy for TSLM with knot
TSLM__dynamic_accuracy_knot <- forecast_TSLM_dynamic_knot %>%
  accuracy(CAN_transform_test)

TSLM__dynamic_accuracy_knot

# Plot the forecasts along with the original data for TSLM with knot
forecast_TSLM_dynamic_knot %>%
  autoplot(CAN_transform, level = NULL) +
  labs(
    title = "Forecasts for TSLM (dynamic) Model with Knot at 2021",
    y = "Value",
    x = "Year"
  ) +
  theme_minimal()
```
## (g) Hybrid TSLM_external with knot + ETS
```{r}
# Step 1: Fit TSLM model with knot at 2021
tslm_fit <- CAN_transform_train %>%
  model(tslm = TSLM(Immigrants ~ Deaths + Births + ReEmigrants + trend(knots = 2021)))

# Step 2: Extract residuals and prepare tsibble
residuals_tslm <- augment(tslm_fit) %>%
  select(Year, residuals = .resid) %>% 
  as_tsibble(index = Year)

residuals_tslm

# Step 3: Fit ETS model on residuals
ets_residuals <- residuals_tslm %>%
  model(ETS = ETS(residuals))

# Step 4: Forecast ETS residuals
forecast_ets_residuals <- ets_residuals %>%
  forecast(h = nrow(CAN_transform_test))

# Step 5: Forecast TSLM with knot on the test set
forecast_TSLM <- tslm_fit %>%
  forecast(new_data = CAN_transform_test)

# Step 6: Combine TSLM predictions with ETS residual forecasts
combined_forecast <- forecast_TSLM %>%
  mutate(.mean = .mean + forecast_ets_residuals$.mean)

# Step 7: Evaluate combined forecast accuracy
accuracy_combined <- combined_forecast %>%
  accuracy(CAN_transform_test)

# Print results
print(accuracy_combined)

# Step 8: Plot combined forecasts
combined_forecast %>%
  autoplot(CAN_transform, level = NULL) +
  labs(
    title = "Forecasts for TSLM with Knot + ETS Residuals",
    y = "Immigrants",
    x = "Year"
  ) +
  theme_minimal()

```
# 7. Deep learning models
```{r}
library(vars)
library(tseries)
```
## (a) VAR
```{r}
# Select optimal lag
lag_selection <- VARselect(CAN_transform, lag.max = 5, type = "const")
optimal_lag <- lag_selection$selection["AIC(n)"]

# Perform PCA
pca_result <- prcomp(CAN_transform_train, scale. = TRUE)

# Use the principal components as input to VAR
CAN_pca <- pca_result$x[, 1:4]  # Select the top 4 principal components
var_model_pca <- VAR(CAN_pca, p = optimal_lag, type = "const")

# Check residuals
library(vars)
serial.test(var_model_pca, lags.pt = 10, type = "PT.asymptotic")
normality.test(var_model_pca)

# Forecast for test period
forecast_steps <- nrow(CAN_transform_test)
var_forecast <- predict(var_model_pca, n.ahead = forecast_steps)

pc1_forecast <- var_forecast$fcst$PC1[, "fcst"] 
pc1_forecast

# Step 2: Perform PCA on CAN_transform_test to get actual PC1 values
test_pca <- predict(pca_result, newdata = as.matrix(CAN_transform_test %>% select(-Year)))
actual_pc1 <- test_pca[, 1] 

actual_pc1
```
```{r}
# Combine into a data frame for evaluation
comparison_df <- data.frame(
  Year = CAN_transform_test$Year,
  Forecasted_PC1 = pc1_forecast,
  Actual_PC1 = actual_pc1
)

# Step 4: Calculate Accuracy Metrics
# Mean Absolute Error (MAE) and Root Mean Square Error (RMSE)
mae <- mean(abs(comparison_df$Forecasted_PC1 - comparison_df$Actual_PC1))
rmse <- sqrt(mean((comparison_df$Forecasted_PC1 - comparison_df$Actual_PC1)^2))

# Print the results
cat("Accuracy Metrics for PC1:\n")
cat("Mean Absolute Error (MAE):", mae, "\n")
cat("Root Mean Square Error (RMSE):", rmse, "\n")

# Step 5: Visualize the forecasted vs. actual PC1
library(ggplot2)
ggplot(comparison_df, aes(x = Year)) +
  geom_line(aes(y = Forecasted_PC1, color = "Forecasted"), size = 1.2) +
  geom_line(aes(y = Actual_PC1, color = "Actual"), size = 1.2) +
  labs(
    title = "Comparison of Forecasted vs. Actual PC1",
    y = "PC1 Value",
    x = "Year"
  ) +
  scale_color_manual(values = c("Forecasted" = "blue", "Actual" = "red")) +
  theme_minimal()
```

## (b) XGBoost
```{r}
library(xgboost)
```
```{r}
# Convert training data into a matrix format suitable for XGBoost
x_train <- CAN_transform_train %>% 
  select(Deaths, Births, ReEmigrants) %>% 
  as.matrix()
y_train <- CAN_transform_train$Immigrants

# Convert test data into matrix format
x_test <- CAN_transform_test %>%
  select(Deaths, Births, ReEmigrants) %>%
  as.matrix()

# Set parameters for XGBoost
params <- list(
  booster = "gbtree",
  objective = "reg:squarederror", # Regression problem
  eta = 0.1,                      # Learning rate
  max_depth = 6,                  # Maximum depth of trees
  subsample = 0.8,                # Subsample ratio of training data
  colsample_bytree = 0.8          # Subsample ratio of columns
)

# Train the XGBoost model
xgb_model <- xgboost(
  data = x_train, 
  label = y_train, 
  params = params, 
  nrounds = 100,    # Number of boosting rounds
  verbose = 0       # Suppress training output
)

# Make predictions on the test set
xgb_predictions <- predict(xgb_model, x_test)

# Evaluate model performance
y_test <- CAN_transform_test$Immigrants

# Calculate Mean Absolute Error (MAE)
mae <- mean(abs(y_test - xgb_predictions))

# Calculate Root Mean Squared Error (RMSE)
rmse <- sqrt(mean((y_test - xgb_predictions)^2))

# Calculate Mean Absolute Percentage Error (MAPE)
mape <- mean(abs((y_test - xgb_predictions) / y_test)) * 100

# Print metrics
cat("Evaluation Metrics:\n")
cat("Mean Absolute Error (MAE):", mae, "\n")
cat("Root Mean Squared Error (RMSE):", rmse, "\n")
cat("Mean Absolute Percentage Error (MAPE):", mape, "%\n")


```
```{r}
# Visualize predictions vs actual values
library(ggplot2)
data.frame(
  Actual = y_test,
  Predicted = xgb_predictions
) %>%
  ggplot(aes(x = Actual, y = Predicted)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  labs(title = "XGBoost Predictions vs Actual",
       x = "Actual Values",
       y = "Predicted Values") +
  theme_minimal()
```
## (c) XGBoost for TS

```{r}
# Add lagged features for Immigrants
CAN_transform_train <- CAN_transform_train %>%
  mutate(
    Immigrants_Lag1 = dplyr::lag(Immigrants, 1),
    Immigrants_Lag2 = dplyr::lag(Immigrants, 2)
  ) %>%
  drop_na() 

# Split features and target
x_train <- CAN_transform_train %>%
  select(Immigrants_Lag1, Immigrants_Lag2, Deaths, Births, ReEmigrants) %>%
  as.matrix()

y_train <- CAN_transform_train$Immigrants

# Create DMatrix for training
dtrain <- xgb.DMatrix(data = x_train, label = y_train)

# Set XGBoost parameters
params <- list(
  objective = "reg:squarederror", # Regression objective
  eval_metric = "rmse",          # Metric to evaluate model performance
  eta = 0.1,                     # Learning rate
  max_depth = 6,                 # Depth of trees
  subsample = 0.8,               # Row sampling
  colsample_bytree = 0.8         # Column sampling
)

# Train XGBoost model
xgb_model <- xgb.train(
  params = params,
  data = dtrain,
  nrounds = 100, # Number of boosting rounds
  verbose = 1
)

```

```{r}
# Prepare test data with lagged features
CAN_transform_test <- CAN_transform_test %>%
  mutate(
    Immigrants_Lag1 = dplyr::lag(Immigrants, 1),
    Immigrants_Lag2 = dplyr::lag(Immigrants, 2)
  ) %>%
  drop_na()

x_test <- CAN_transform_test %>%
  select(Immigrants_Lag1, Immigrants_Lag2, Deaths, Births, ReEmigrants) %>%
  as.matrix()

y_test <- CAN_transform_test$Immigrants

# Make predictions
y_pred <- predict(xgb_model, x_test)

# Evaluate Model Performance

# Calculate RMSE
rmse <- sqrt(mean((y_test - y_pred)^2))

# Calculate MAPE
mape <- mean(abs((y_test - y_pred) / y_test)) * 100

# Calculate MAE
mae <- mean(abs(y_test - y_pred))

# Print results
cat("RMSE:", rmse, "\n")
cat("MAE:", mae, "\n")
cat("MAPE:", mape, "%\n")
```
```{r}
# Combine actual and predicted values
results <- data.frame(
  Actual = y_test,
  Predicted = y_pred,
  Time = seq_along(y_test)
)

# Plot
ggplot(results, aes(x = Time)) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Predicted, color = "Predicted")) +
  labs(title = "Actual vs Predicted Values",
       y = "Immigrants",
       color = "Legend") +
  theme_minimal()
```
## (d) Hybrid ARIMA - XGBoost
```{r}
# Fit ARIMA model
arima_model <- auto.arima(
  CAN_transform_train$Immigrants, 
  xreg = CAN_transform_train %>% select(Deaths, Births, ReEmigrants) %>% as.matrix()
)

# Forecast with ARIMA
arima_forecast <- forecast(
  arima_model, 
  xreg = CAN_transform_test %>% select(Deaths, Births, ReEmigrants) %>% as.matrix()
)

# Extract residuals
arima_residuals <- residuals(arima_model)
```

```{r}
# Add lagged features and residuals for XGBoost
CAN_transform_train <- CAN_transform_train %>%
  mutate(
    Residuals = arima_residuals,
    Immigrants_Lag1 = dplyr::lag(Immigrants, 1),
    Immigrants_Lag2 = dplyr::lag(Immigrants, 2)
  ) %>%
  drop_na() # Remove rows with NA

# Create training matrix for XGBoost
x_train <- CAN_transform_train %>%
  select(Immigrants_Lag1, Immigrants_Lag2, Deaths, Births, ReEmigrants) %>%
  as.matrix()

y_train <- CAN_transform_train$Residuals
```

```{r}
# Create DMatrix for training
dtrain <- xgb.DMatrix(data = x_train, label = y_train)

# Set XGBoost parameters
params <- list(
  objective = "reg:squarederror",
  eval_metric = "rmse",
  eta = 0.1,
  max_depth = 6,
  subsample = 0.8,
  colsample_bytree = 0.8
)

# Train XGBoost model
xgb_model <- xgb.train(
  params = params,
  data = dtrain,
  nrounds = 100,
  verbose = 1
)
```

```{r warning = FALSE}
# Prepare test data for XGBoost
CAN_transform_test <- CAN_transform_test %>%
  mutate(
    Immigrants_Lag1 = dplyr::lag(Immigrants, 1),
    Immigrants_Lag2 = dplyr::lag(Immigrants, 2)
  ) %>%
  drop_na()

x_test <- CAN_transform_test %>%
  select(Immigrants_Lag1, Immigrants_Lag2, Deaths, Births, ReEmigrants) %>%
  as.matrix()

# Predict residuals with XGBoost
xgb_residuals <- predict(xgb_model, x_test)

# Combine ARIMA and XGBoost predictions
final_predictions <- arima_forecast$mean + xgb_residuals
```

```{r warning = FALSE}
# Actual values
y_test <- CAN_transform_test$Immigrants

# Calculate RMSE
rmse <- sqrt(mean((y_test - final_predictions)^2))

# Calculate MAPE
mape <- mean(abs((y_test - final_predictions) / y_test)) * 100

# Calculate MAE
mae <- mean(abs(y_test - final_predictions))

# Print results
cat("Hybrid Model RMSE:", rmse, "\n")
cat("Hybrid Model MAE:", mae, "\n")
cat("Hybrid Model MAPE:", mape, "%\n")

```
```{r}
# Combine results
results <- data.frame(
  Actual = y_test,
  Predicted = final_predictions,
  Time = seq_along(y_test)
)

# Plot
ggplot(results, aes(x = Time)) +
  geom_line(aes(y = Actual, color = "Actual")) +
  geom_line(aes(y = Predicted, color = "Predicted")) +
  labs(title = "Hybrid Model: Actual vs Predicted",
       y = "Immigrants",
       color = "Legend") +
  theme_minimal()
```

## (e) XGBoost + ARIMA
```{r}
# Prepare features for XGBoost (using lags and explanatory variables)
CAN_transform_train <- CAN_transform_train %>%
  mutate(
    Immigrants_Lag1 = dplyr::lag(Immigrants, 1),
    Immigrants_Lag2 = dplyr::lag(Immigrants, 2)
  ) %>%
  drop_na()

x_train <- CAN_transform_train %>%
  select(Immigrants_Lag1, Immigrants_Lag2, Deaths, Births, ReEmigrants) %>%
  as.matrix()

y_train <- CAN_transform_train$Immigrants  # Target is the Immigration values

# Create DMatrix for XGBoost
dtrain <- xgb.DMatrix(data = x_train, label = y_train)

# Set XGBoost parameters
params <- list(
  objective = "reg:squarederror",
  eval_metric = "rmse",
  eta = 0.1,
  max_depth = 6,
  subsample = 0.8,
  colsample_bytree = 0.8
)

# Train XGBoost model
xgb_model <- xgb.train(
  params = params,
  data = dtrain,
  nrounds = 100,
  verbose = 1
)

```

```{r}
# Prepare test data for XGBoost
CAN_transform_test <- CAN_transform_test %>%
  mutate(
    Immigrants_Lag1 = dplyr::lag(Immigrants, 1),
    Immigrants_Lag2 = dplyr::lag(Immigrants, 2)
  ) %>%
  drop_na()

x_test <- CAN_transform_test %>%
  select(Immigrants_Lag1, Immigrants_Lag2, Deaths, Births, ReEmigrants) %>%
  as.matrix()

# Predict immigration values using XGBoost
xgb_predictions <- predict(xgb_model, x_test)

# Calculate residuals from XGBoost predictions
xgb_residuals <- CAN_transform_test$Immigrants - xgb_predictions

# Fit ARIMA model on residuals
arima_model <- auto.arima(xgb_residuals)

# Forecast with ARIMA
arima_forecast <- forecast(arima_model, h = length(xgb_residuals))

# Combine XGBoost predictions with ARIMA forecast on residuals
final_predictions <- xgb_predictions + arima_forecast$mean


```
```{r}
# Actual immigration values
y_test <- CAN_transform_test$Immigrants

# Calculate RMSE
rmse <- sqrt(mean((y_test - final_predictions)^2))

# Calculate MAPE
mape <- mean(abs((y_test - final_predictions) / y_test)) * 100

# Calculate MAE
mae <- mean(abs(y_test - final_predictions))

# Print results
cat("Hybrid Model RMSE:", rmse, "\n")
cat("Hybrid Model MAE:", mae, "\n")
cat("Hybrid Model MAPE:", mape, "%\n")

```






