---
title: "Group Project"
author: "Thai Pham - T00727094"
date: "2024-11-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import necessary libraries

```{r install_packages, echo = FALSE, message = FALSE, warning = FALSE}
pkg_list = c("dplyr", "ggplot2", "tidyr", "readr", "fable", "feasts", "fpp3", "lubridate", 'fabletools',
             "purrr", "tibble", "stringr", "forcats", 'tsibble', "rugarch", "rsample", 
             "lubridate", "tsibble", "forecast", "caret", 'GGally', "ggcorrplot", "Metrics",
             "randomForest", "e1071", "xgboost", "glmnet", "quantmod", "gridExtra", "lubridate", "tseries", "rsample", "prophet")

# Install and load required packages
for (pkg in pkg_list) {
  # Check if the package is installed
  if (!require(pkg, character.only = TRUE)) {
    # Install if not available
    install.packages(pkg, dependencies = TRUE)
    # Load the package after installation
    library(pkg, character.only = TRUE)
  }
}

```

# Loading data
```{r}
# Step 1: Load the data
data <- read.csv("CANSIM.csv")

# Display the column names of the dataframe 
colnames(data)
unique(data$Components.of.population.growth)

# Step 2: Create tsibble data
Cansim <- data %>%
  mutate(
    Year = as.integer(substr(REF_DATE, 1, 4))  # Extract the first year as integer
  ) %>%
  select(
    Year,                      
    Geo = GEO,                 
    Components = Components.of.population.growth,  
    Value = VALUE             
  )

Cansim <- as_tsibble(Cansim, index = Year, key = c(Geo, Components))

# View the data
head(Cansim)

# Extract the data of interest
Can_Immigrant <- Cansim %>% filter (Geo == "Canada", Components == "Immigrants")

Can_Immigrant$Adjust_Value <- Can_Immigrant$Value / 10^5
tail(Can_Immigrant)
``` 

# 1. EDA
```{r}
# Check for missing values across all columns
colSums(is.na(Can_Immigrant))

# Plot the data
Can_Immigrant %>% autoplot(Value)
```
## Box-cox transformation
```{r}
# Apply Box-Cox transformation using Guerrero method
lambda <- BoxCox.lambda(Can_Immigrant$Value, method = "guerrero")  
Can_Immigrant <- Can_Immigrant %>% mutate(boxcox_Value = BoxCox(Value, lambda))
lambda
tail(Can_Immigrant)

# Plot original vs Box-Cox transformed series
autoplot(Can_Immigrant, Value) + 
  ggtitle("Original immigrants") +
  labs(y = "People")

autoplot(Can_Immigrant, boxcox_Value) + 
  ggtitle("Box-Cox Transformed Immigrants (Guerrero)") +
  labs(y = "Box-Cox Transformed Immigrants")

# ACF Plot
Can_Immigrant %>% ACF(Value) %>%
  autoplot() + labs(subtitle = "ACF plot of original data")

# PACF Plot
Can_Immigrant %>% PACF(Value) %>%
  autoplot() + labs(subtitle = "PACF plot of original data")
```

## Stationary and differencing check for original data
```{r}
# Check the stationarity of TS using KPSS test
Can_Immigrant |>
  features(Value, unitroot_kpss)
# Way 2 of using KPSS
KPSSTest = kpss.test(Can_Immigrant$Value)
KPSSTest

# Check the stationarity of TS using Augmented Dickey-Fuller (ADF) test
adf_test <- adf.test(Can_Immigrant$Value, alternative = "stationary")
print(adf_test)

# Determine whether differencing is required
Can_Immigrant |>
  features(Value, unitroot_ndiffs)

# Determine whether seasonal differencing is required
Can_Immigrant |>
  features(Value, unitroot_nsdiffs)
```
*KPSS test*: the p-value is shown as 0.01 (and therefore it may be smaller than that), indicating that the null hypothesis is rejected. That is, the data are not stationary. 
*ADF test*: the p-value = 0.4518 (larger than 0.05), indicating that the null hypothesis is NOT rejected. That is, the data are not stationary. 

## Stationary and differencing check for box-cox data
```{r}
# Check the stationarity of TS using KPSS test
Can_Immigrant |>
  features(boxcox_Value, unitroot_kpss)
# Way 2 of using KPSS
KPSSTest = kpss.test(Can_Immigrant$Value)
KPSSTest

# Check the stationarity of TS using Augmented Dickey-Fuller (ADF) test
adf_test <- adf.test(Can_Immigrant$boxcox_Value, alternative = "stationary")
print(adf_test)

# Determine whether differencing is required
Can_Immigrant |>
  features(boxcox_Value, unitroot_ndiffs)

# Determine whether seasonal differencing is required
Can_Immigrant |>
  features(boxcox_Value, unitroot_nsdiffs)
```
## Do the first differencing on box-cox data and check again
```{r}
# Do the first differencing and check again
Can_Immigrant <- Can_Immigrant %>%
  mutate(diff_Value = difference(boxcox_Value)) 

Can_Immigrant|>
  features(diff_Value, unitroot_kpss)

Can_Immigrant

```
This time, the p-value is reported as 0.1 (and so it could be larger than that). We can conclude that the differenced data appear stationary.

## Plot the original data with transformed and differencing data
```{r}
Can_Immigrant |>
  transmute(
    `Value` = Value,  
    `adjusted_Value` = boxcox_Value,
    `Annual change in adjusted Value` = diff_Value
  ) |>
  pivot_longer(-Year, names_to = "Type", values_to = "Value") |>
  mutate(
    Type = factor(Type, levels = c(
      "Value",
      "adjusted_Value",
      "Annual change in adjusted Value"
    ))
  ) |>
  as_tsibble(index = Year) |>
  ggplot(aes(x = Year, y = Value)) +
  geom_line() +
  facet_grid(vars(Type), scales = "free_y") +
  labs(title = "Canadian Statistic of Immigrants Analysis", y = NULL)

```
# 2. Split data into train and test sets
```{r}
# Generate train and test data
Can_Immigrant_train <- Can_Immigrant %>% filter(Year <= 2018)
Can_Immigrant_test <- Can_Immigrant %>% filter(Year > 2018)

Can_Immigrant_train
# Plot train and test data
Can_Immigrant_train %>%
  autoplot(Value) +
  autolayer(Can_Immigrant_test, Value, colour = "red") +
  labs(title = "Train and Test Data", y = "adjusted_Value")
```
# 3. Set up time series cross-validation 
```{r cache = TRUE}
Can_Immigrant_train_tr <- Can_Immigrant_train |>
stretch_tsibble(.init = 5, .step = 1) 
```
# 4. Fit benchmark models
## Fit model
```{r}
benchmark_fit <- Can_Immigrant_train %>%
   model(
    Naive = NAIVE(diff_Value),                   #Naive model
    Drift = RW(diff_Value ~ drift()),            #Drift model
    Mean = MEAN(diff_Value),                     #Mean model
    RandomWalk= RW(diff_Value)                   #Random walk model              
  )
```

## Residuals diagnostics
```{r warning = FALSE}
# Check residuals for Naive model
gg_tsresiduals(benchmark_fit %>% select(Naive))

# For Drift model
gg_tsresiduals(benchmark_fit %>% select(Drift))

# For Mean model
gg_tsresiduals(benchmark_fit %>% select(Mean)) 

# For Random Walk model
gg_tsresiduals(benchmark_fit %>% select(RandomWalk))
```
## Portmanteau test
```{r}
augment(benchmark_fit) %>% features(.resid, ljung_box, lag=10, dof=0)
```
None of the benchmark models pass the Ljung-box test!

## Forecast and evaluate accuracy
```{r cache = TRUE}
# Time series cross-validation accuracy
cv_bench_forecast<- Can_Immigrant_train_tr %>%
  model(
    Naive = NAIVE(diff_Value),                   #Naive model
    Drift = RW(diff_Value ~ drift()),            #Drift model
    Mean = MEAN(diff_Value),                     #Mean model
    RandomWalk= RW(diff_Value)                   #Random walk model              
  ) %>% 
  forecast(h = 5) 
accuracy(cv_bench_forecast, Can_Immigrant_train )

```
```{r}
# Produce forecast
benchmark_models <- Can_Immigrant_train %>%
  model(
    Naive = NAIVE(diff_Value),                   # Naive model
    Drift = RW(diff_Value ~ drift()),            # Drift model
    Mean = MEAN(diff_Value),                     # Mean model
    RandomWalk = RW(diff_Value)                  # Random walk model
  )

# Produce forecasts
forecast_benchmark <- benchmark_models %>%
  forecast(new_data = Can_Immigrant_test)

# Evaluate forecast accuracy
benchmark_accuracy <- forecast_benchmark %>%
  accuracy(Can_Immigrant_test)

benchmark_accuracy 

# Plot the forecasts along with the original data
forecast_benchmark  %>%
  autoplot(Can_Immigrant, level = NULL) +
  labs(
    title = "Forecasts for Benchmark Models",
    y = "Value",
    x = "Year"
  ) +
  theme_minimal()
```
## Calculate accuracy manually
```{r}
# Step 1: Prepare the time series data
ts_data <- CAN_transform %>%
  select(Year, Immigrants) %>%
  as_tsibble(index = Year)

# Step 2: Function to calculate RMSE, MAE, and MAPE
calculate_metrics <- function(actual, predicted) {
  rmse <- sqrt(mean((actual - predicted)^2, na.rm = TRUE))
  mae <- mean(abs(actual - predicted), na.rm = TRUE)
  mape <- mean(abs((actual - predicted) / actual), na.rm = TRUE) * 100
  return(c(RMSE = rmse, MAE = mae, MAPE = mape))
}

# Step 3: Fit models and calculate metrics for each model
models <- c("Mean", "Naive", "SeasonalNaive", "Drift")
results <- data.frame(Model = models, RMSE = NA, MAE = NA, MAPE = NA)

# Prepare the time series object for forecasting
ts_series <- ts(ts_data$Immigrants, start = min(ts_data$Year), frequency = 1)

# Create an empty list to store forecasted values for plotting
forecast_data <- list()

for (model in models) {
  # Fit the model using tsCV and calculate errors
  if (model == "Mean") {
    model_fit <- meanf(ts_series, h = 1)
  } else if (model == "Naive") {
    model_fit <- naive(ts_series, h = 1)
  } else if (model == "SeasonalNaive") {
    model_fit <- snaive(ts_series, h = 1)
  } else if (model == "Drift") {
    model_fit <- rwf(ts_series, drift = TRUE, h = 1)
  }
  
  # Cross-validation using tsCV
  cv_errors <- tsCV(ts_series, forecastfunction = function(x, h) forecast(x, h = h), h = 1)
  
  # Ensure the lengths match for predicted values and actual values
  predicted_values <- forecast(model_fit, h = 1)$mean
  
  # Extract the actual values for the corresponding time points
  actual_values <- ts_series[(length(ts_series) - length(predicted_values) + 1):length(ts_series)]
  
  # Calculate RMSE, MAE, and MAPE for the current model
  metrics <- calculate_metrics(actual_values, predicted_values)
  
  # Store the results
  results[results$Model == model, 2:4] <- metrics
}

# Step 4: Display the results as a table
print(results)

```

# 5. Fit other TS models
## ARIMA
```{r}
# Fit ARIMA model 
ARIMA_fit <- Can_Immigrant_train %>%
   model(Arima = ARIMA(Value)
  )

# Check residuals for ARIMA model
gg_tsresiduals(ARIMA_fit)

# Portmanteau test
augment(ARIMA_fit) %>% features(.resid, ljung_box, lag=10, dof=0)

# Time series cross-validation accuracy
Can_Immigrant_train_tr %>%
  model (Arima = ARIMA(Value)) %>% 
  forecast(h = 5) %>%
  accuracy(Can_Immigrant_train)

# Produce forecasts
forecast_ARIMA <- ARIMA_fit %>%
  forecast(new_data = Can_Immigrant_test)

# Evaluate forecast accuracy
ARIMA_accuracy <- forecast_ARIMA %>%
  accuracy(Can_Immigrant_test)

ARIMA_accuracy

# Plot the forecasts along with the original data
forecast_ARIMA  %>%
  autoplot(Can_Immigrant, level = NULL) +
  labs(
    title = "Forecasts for ARIMA Model",
    y = "Value",
    x = "Year"
  ) +
  theme_minimal()
```

## NNAR
```{r cache = TRUE}
# Fit NNAR model 
NNAR_fit <- Can_Immigrant_train %>%
   model(NNETAR(Value))

# Check residuals for ARIMA model
gg_tsresiduals(NNAR_fit)

# Portmanteau test
augment(NNAR_fit) %>% features(.resid, ljung_box, lag=10, dof=0)

# Produce forecasts
forecast_NNAR <- NNAR_fit %>%
  forecast(new_data = Can_Immigrant_test)

# Evaluate forecast accuracy
NNAR_accuracy <- forecast_NNAR %>%
  accuracy(Can_Immigrant_test)

NNAR_accuracy

# Plot the forecasts along with the original data
forecast_NNAR  %>%
  autoplot(Can_Immigrant, level = NULL) +
  labs(
    title = "Forecasts for NNAR Model",
    y = "Value",
    x = "Year"
  ) +
  theme_minimal()
```
## TSLM
```{r}
# Fit TSLM model
TSLM_fit <- Can_Immigrant_train %>%
  model(TSLM = TSLM(Value~trend()))

# Check residuals for TSLM model
gg_tsresiduals(TSLM_fit)

# Portmanteau test for TSLM model residuals
augment(TSLM_fit) %>% features(.resid, ljung_box, lag = 10, dof = 0)

# Time series cross-validation accuracy for TSLM
Can_Immigrant_train_tr %>%
  model(TSLM = TSLM(Value ~ trend() + season())) %>%
  forecast(h = 5) %>%
  accuracy(Can_Immigrant_train)

# Produce forecasts with TSLM
forecast_TSLM <- TSLM_fit %>%
  forecast(new_data = Can_Immigrant_test)

# Evaluate forecast accuracy for TSLM
TSLM_accuracy <- forecast_TSLM %>%
  accuracy(Can_Immigrant_test)

TSLM_accuracy

# Plot the forecasts along with the original data for TSLM
forecast_TSLM %>%
  autoplot(Can_Immigrant, level = NULL) +
  labs(
    title = "Forecasts for TSLM Model",
    y = "Value",
    x = "Year"
  ) +
  theme_minimal()
```
**TSLM WITH DUMMY VARIABLE**

```{r}
# Dummy variable for year 2021 due to TR2PR event
Can_Immigrant_train <- Can_Immigrant_train %>%
  mutate(TR2PR = ifelse(Year >= 2021, 1, 0))

Can_Immigrant_test <- Can_Immigrant_test %>%
  mutate(TR2PR = ifelse(Year >= 2021, 1, 0))

# Fit TSLM model with dummy variable
TSLM_fit_dummy <- Can_Immigrant_train %>%
  model(TSLM = TSLM(Value ~ trend() + TR2PR))

report(TSLM_fit_dummy)

# Check residuals for TSLM model with dummy
gg_tsresiduals(TSLM_fit_dummy)

# Portmanteau test for TSLM model residuals
augment(TSLM_fit_dummy) %>% features(.resid, ljung_box, lag = 10, dof = 0)

# Produce forecasts with TSLM with dummy
forecast_TSLM_dummy <- TSLM_fit_dummy %>%
  forecast(new_data = Can_Immigrant_test)

# Evaluate forecast accuracy for TSLM with dummy
TSLM_accuracy_dummy <- forecast_TSLM_dummy %>%
  accuracy(Can_Immigrant_test)

TSLM_accuracy_dummy

# Plot the forecasts along with the original data for TSLM with dummy
forecast_TSLM_dummy %>%
  autoplot(Can_Immigrant, level = NULL) +
  labs(
    title = "Forecasts for TSLM Model with Dummy Variable",
    y = "Value",
    x = "Year"
  ) +
  theme_minimal()

```
**TSLM WITH KNOT**
```{r}
# Fit TSLM model with knot at year 2021
TSLM_fit_knot <- Can_Immigrant_train %>%
  model(piecewise = TSLM(Value ~ trend(knots = 2021))
  )
 

# Check residuals for TSLM model with knot
gg_tsresiduals(TSLM_fit_knot)

# Portmanteau test for TSLM model residuals
augment(TSLM_fit_knot) %>% features(.resid, ljung_box, lag = 10, dof = 0)

# Time series cross-validation accuracy for TSLM with knot
Can_Immigrant_train_tr %>%
  model(piecewise = TSLM(Value ~ trend(knots = 2021))) %>%
  forecast(h = 5) %>%
  accuracy(Can_Immigrant_train)

# Produce forecasts with TSLM with knot
forecast_TSLM_knot <- TSLM_fit_knot %>%
  forecast(new_data = Can_Immigrant_test)

# Evaluate forecast accuracy for TSLM with knot
TSLM_accuracy_knot <- forecast_TSLM_knot %>%
  accuracy(Can_Immigrant_test)

TSLM_accuracy_knot

# Plot the forecasts along with the original data for TSLM with knot
forecast_TSLM_knot %>%
  autoplot(Can_Immigrant, level = NULL) +
  labs(
    title = "Forecasts for TSLM Model with Knot at 2021",
    y = "Value",
    x = "Year"
  ) +
  theme_minimal()
```

## Prophet
```{r}
# Prepare the data for Prophet
Can_Immigrant_train_prophet <- Can_Immigrant_train %>%
  rename(ds = Year, y = Value)

Can_Immigrant_train_prophet$ds <- as.Date(paste(Can_Immigrant_train_prophet$ds, "01-01", sep = "-"), format = "%Y-%m-%d")


# Fit the Prophet model
prophet_fit <- prophet(Can_Immigrant_train_prophet, yearly.seasonality = TRUE, weekly.seasonality = FALSE, daily.seasonality = FALSE)

# Forecast using the Prophet model
forecast_prophet <- predict(prophet_fit, Can_Immigrant_train_prophet)

```

```{r}
# Prepare the test data for Prophet
Can_Immigrant_test_prophet <- Can_Immigrant_test %>%
  rename(ds = Year, y = Value)

Can_Immigrant_test_prophet$ds <- as.Date(paste(Can_Immigrant_test_prophet$ds, "01-01", sep = "-"), format = "%Y-%m-%d")

# Forecast for the test dataset
forecast_test_prophet <- predict(prophet_fit, Can_Immigrant_test_prophet)

# Calculate the residuals (differences between actual and predicted values)
residuals_test_prophet <- Can_Immigrant_test_prophet$y - forecast_test_prophet$yhat

# Calculate accuracy metrics
mae <- mean(abs(residuals_test_prophet))  # Mean Absolute Error
mse <- mean(residuals_test_prophet^2)     # Mean Squared Error
rmse <- sqrt(mse)                         # Root Mean Squared Error

# Print the accuracy metrics
cat("MAE:", mae, "\n")
cat("MSE:", mse, "\n")
cat("RMSE:", rmse, "\n")

```
```{r}
# Combine the forecast with the original data for plotting
original_data <- Can_Immigrant_train_prophet %>%
  bind_rows(Can_Immigrant_test_prophet) %>%
  mutate(forecast = c(forecast_prophet$yhat, forecast_test_prophet$yhat))

# Plot the original data and the forecasted data
ggplot() +
  geom_line(data = original_data, aes(x = ds, y = y), color = "blue", size = 1, linetype = "dashed") +  # Original data (blue)
  geom_line(data = original_data, aes(x = ds, y = forecast), color = "red", size = 1) +  # Forecasted data (red)
  labs(
    title = "Forecast vs Original Data for PROPHET model",
    x = "Year",
    y = "Value"
  ) +
  theme_minimal()
```
## ETS
```{r}
# Prepare the data for ETS
Can_Immigrant_train_ts <- ts(Can_Immigrant_train$Value, start = c(min(Can_Immigrant_train$Year)), frequency = 1)

# Fit the ETS model
ets_fit <- ets(Can_Immigrant_train_ts)

# Check residuals for the ETS model
checkresiduals(ets_fit)

# Compute accuracy on the training data
accuracy_train <- accuracy(ets_fit)

# Produce forecasts for the future period (length of the test data)
forecast_ets <- forecast(ets_fit, h = length(Can_Immigrant_test$Value))

# Evaluate forecast accuracy on the test set
ets_accuracy <- accuracy(forecast_ets, Can_Immigrant_test$Value)

# Print forecast accuracy
print(ets_accuracy)

# Plot the forecasts along with the original data
autoplot(forecast_ets) +
  geom_line(data = Can_Immigrant_train, aes(x = Year, y = Value), color = "blue", size = 1) +
  labs(
    title = "Forecasts for ETS Model",
    y = "Value",
    x = "Year"
  ) +
  theme_minimal()
```
## EWMA
```{r}
# Define EWMA function
EWMA_forecast <- function(data, alpha, forecast_horizon) {
  n <- length(data)
  smoothed_values <- numeric(n)
  smoothed_values[1] <- data[1] # Initialize with the first value
  
  # Calculate EWMA
  for (i in 2:n) {
    smoothed_values[i] <- alpha * data[i] + (1 - alpha) * smoothed_values[i - 1]
  }
  
  # Forecast future values using the last smoothed value
  future_values <- rep(smoothed_values[n], forecast_horizon)
  
  list(smoothed = smoothed_values, forecast = future_values)
}

# Define a range of alpha values to test
alpha_values <- seq(0.1, 0.9, by = 0.1)

# Initialize an empty list to store results
results <- tibble()

# Initialize an empty tibble with the required columns
results <- tibble(
  alpha = numeric(),
  MAE = numeric(),
  RMSE = numeric()
)

# Loop through alpha values to calculate accuracy
for (alpha in alpha_values) {
  # Apply EWMA with the current alpha
  EWMA_result <- EWMA_forecast(Can_Immigrant_train$Value, alpha, length(Can_Immigrant_test$Value))
  
  # Calculate accuracy for the test set
  actual_test <- Can_Immigrant_test$Value
  forecast_test <- EWMA_result$forecast
  
  MAE <- mean(abs(actual_test - forecast_test), na.rm = TRUE)
  RMSE <- sqrt(mean((actual_test - forecast_test)^2, na.rm = TRUE))
  
  # Add results for the current alpha to the tibble
  results <- results %>%
    add_row(alpha = alpha, MAE = MAE, RMSE = RMSE)
}

# Find the best alpha based on MAE or RMSE
best_alpha <- results %>% filter(MAE == min(MAE)) %>% pull(alpha)

# Print the results
print(results)

# Apply the best alpha for final forecast
EWMA_best <- EWMA_forecast(Can_Immigrant_train$Value, best_alpha, length(Can_Immigrant_test$Value))

# Combine original data and forecasts for plotting
EWMA_forecasted <- tibble(
  Year = c(Can_Immigrant_train$Year, Can_Immigrant_test$Year),
  Value = c(Can_Immigrant_train$Value, rep(NA, length(Can_Immigrant_test$Value))),
  EWMA = c(EWMA_best$smoothed, EWMA_best$forecast)
)

# Plot the EWMA results with the best alpha
EWMA_forecasted %>%
  ggplot(aes(x = Year)) +
  geom_line(aes(y = Value), color = "blue", size = 1.2, linetype = "dashed", na.rm = TRUE) +
  geom_line(aes(y = EWMA), color = "red", size = 1.2) +
  labs(
    title = paste("EWMA Forecasting (Best Alpha =", best_alpha, ")"),
    y = "Value",
    x = "Year"
  ) +
  theme_minimal()
```
#---------------------------------------------------------------------------------------------------
# IMMIGRANT ANALYSIS IN RELATION TO OTHER INDICATORS
# 1. Preparing data
```{r}
CAN <- data %>%
  mutate(
    Year = as.integer(substr(REF_DATE, 1, 4))
    ) %>%
  select(Year, Components.of.population.growth, VALUE) %>%
  pivot_wider(names_from = Components.of.population.growth, values_from = VALUE) 

# Extract desired columns
CAN_df <- CAN %>% 
  select(Year,Births,Deaths,Immigrants,Emigrants,"Returning emigrants") 

# Create tsibble data
CAN_ts <- CAN %>% 
  select(Year,Births,Deaths,Immigrants,Emigrants,"Returning emigrants") %>%
  as_tsibble(index = Year)

CAN_ts_long <- CAN_ts %>% 
  gather(key = "Variable", value = "Value", -Year)

# Plot the data
ggplot(CAN_ts_long, aes(x = Year, y = Value, color = Variable)) + 
  geom_line() +
  labs(title = "Time Series of Indicators", 
       x = "Year", 
       y = "Value") +
  theme_minimal() +
  theme(legend.title = element_blank())

```
# 2. Statistic analysis
```{r}
# Correlation matrix
cor_matrix <- CAN_df %>%
  select(-Year)%>%
   cor(use = "complete.obs")

# Heatmap plotting
ggcorrplot(cor_matrix, lab = TRUE)
```
# 3. Box-cox transformation and differencing
```{r}
adf.test(CAN_ts$Immigrants, alternative = "stationary")
adf.test(CAN_ts$Births, alternative = "stationary")
adf.test(CAN_ts$Deaths, alternative = "stationary")
adf.test(CAN_ts$Emigrants, alternative = "stationary")
adf.test(CAN_ts$`Returning emigrants`, alternative = "stationary")
```
### a. Immigrants
```{r}
# Apply Box-Cox transformation using Guerrero method
lambda_Immigrant <- BoxCox.lambda(CAN_ts$Immigrants, method = "guerrero")  # Find optimal lambda using Guerrero
CAN_ts <- CAN_ts %>% mutate(boxcox_Immigrants = BoxCox(Immigrants, lambda_Immigrant))

# Determine whether differencing is required
CAN_ts |>
  features(boxcox_Immigrants, unitroot_ndiffs)

# Determine whether seasonal differencing is required
CAN_ts |>
  features(boxcox_Immigrants, unitroot_nsdiffs)

# Do the first differencing
CAN_ts <- CAN_ts %>%  mutate(diff_boxcox_Immigrants = c(NA, diff(boxcox_Immigrants, lag = 1)))

# ADF test on transformed data
CAN_ts_no_na <- CAN_ts %>%
  filter(!is.na(diff_boxcox_Immigrants))

adf.test(CAN_ts_no_na$diff_boxcox_Immigrants, alternative = "stationary")

# KPSS test on transformed data
CAN_ts |>
  features(diff_boxcox_Immigrants, unitroot_kpss)
```
=> For Immigrants variable: do the first differencing on box-cox value.

### b. Births
```{r}
# Apply Box-Cox transformation using Guerrero method
lambda_Births <- BoxCox.lambda(CAN_ts$Births, method = "guerrero")  # Find optimal lambda using Guerrero
CAN_ts <- CAN_ts %>% mutate(boxcox_Births = BoxCox(Births, lambda_Immigrant))

# Determine whether differencing is required
CAN_ts |>
  features(boxcox_Births, unitroot_ndiffs)

# Determine whether seasonal differencing is required
CAN_ts |>
  features(boxcox_Births, unitroot_nsdiffs)

# ADF test on transformed data
adf.test(CAN_ts$boxcox_Births, alternative = "stationary")

# KPSS test on transformed data
CAN_ts |>
  features(boxcox_Births, unitroot_kpss)
```

=> For Births variable: Only do the box-cox transform

### c. Deaths
```{r}
# Apply Box-Cox transformation using Guerrero method
lambda_Deaths <- BoxCox.lambda(CAN_ts$Deaths, method = "guerrero")  # Find optimal lambda using Guerrero
CAN_ts <- CAN_ts %>% mutate(boxcox_Deaths = BoxCox(Deaths, lambda_Deaths))

# Determine whether differencing is required
CAN_ts |>
  features(boxcox_Deaths, unitroot_ndiffs)

# Determine whether seasonal differencing is required
CAN_ts |>
  features(boxcox_Deaths, unitroot_nsdiffs)

# Do the first differencing
CAN_ts <- CAN_ts %>%  mutate(diff_boxcox_Deaths = c(NA, diff(boxcox_Deaths, lag = 1)))

# ADF test on transformed data
CAN_ts_no_na <- CAN_ts %>%
  filter(!is.na(diff_boxcox_Deaths))

adf.test(CAN_ts_no_na$diff_boxcox_Deaths, alternative = "stationary")

# KPSS test on transformed data
CAN_ts |>
  features(diff_boxcox_Deaths, unitroot_kpss)
```
=> For Deaths variable: Do the first differencing on box-cox value.

### d. Emigrants
```{r}
# Apply Box-Cox transformation using Guerrero method
lambda_Emigrants <- BoxCox.lambda(CAN_ts$Emigrants, method = "guerrero")  # Find optimal lambda using Guerrero
CAN_ts <- CAN_ts %>% mutate(boxcox_Emigrants = BoxCox(Emigrants, lambda_Emigrants))

# Determine whether differencing is required
CAN_ts |>
  features(boxcox_Emigrants, unitroot_ndiffs)

# Determine whether seasonal differencing is required
CAN_ts |>
  features(boxcox_Emigrants, unitroot_nsdiffs)

# Do the first differencing
CAN_ts <- CAN_ts %>%  mutate(diff_boxcox_Emigrants = c(NA, diff(boxcox_Emigrants, lag = 1)))

# ADF test on transformed data
CAN_ts_no_na <- CAN_ts %>%
  filter(!is.na(diff_boxcox_Emigrants))

adf.test(CAN_ts_no_na$diff_boxcox_Emigrants, alternative = "stationary")

# KPSS test on transformed data
CAN_ts |>
  features(diff_boxcox_Emigrants, unitroot_kpss)
```
=> For Emigrants variable: Do the first differencing on box-cox value

### e. Returning Emigrants
```{r}
# Apply Box-Cox transformation using Guerrero method
lambda_ReEmigrant <- BoxCox.lambda(CAN_ts$'Returning emigrants', method = "guerrero") 
CAN_ts <- CAN_ts %>% mutate(boxcox_ReEmigrant = BoxCox(Immigrants, lambda_ReEmigrant))

# Determine whether differencing is required
CAN_ts |>
  features(boxcox_ReEmigrant, unitroot_ndiffs)

# Determine whether seasonal differencing is required
CAN_ts |>
  features(boxcox_ReEmigrant, unitroot_nsdiffs)

# Do the first differencing
CAN_ts <- CAN_ts %>%  mutate(diff_boxcox_ReEmigrant = c(NA, diff(boxcox_ReEmigrant, lag = 1)))

# ADF test on transformed data
CAN_ts_no_na <- CAN_ts %>%
  filter(!is.na(diff_boxcox_ReEmigrant))

adf.test(CAN_ts_no_na$diff_boxcox_ReEmigrant, alternative = "stationary")

# KPSS test on transformed data
CAN_ts |>
  features(diff_boxcox_ReEmigrant, unitroot_kpss)
```
=> For Returning Emigrants variable: Do the first differencing on box-cox value

```{r}
CAN_transform <- CAN_ts %>%
  select(Year, boxcox_Births, diff_boxcox_Immigrants, diff_boxcox_Deaths, diff_boxcox_Emigrants, diff_boxcox_ReEmigrant) %>%
  rename(
    Births = boxcox_Births,
    Immigrants = diff_boxcox_Immigrants,
    Deaths = diff_boxcox_Deaths,
    Emigrants = diff_boxcox_Emigrants,
    ReEmigrants = diff_boxcox_ReEmigrant
  ) %>%
  filter (Year > 1971) %>%
  as_tsibble(index = Year)

CAN_transform
```

# 4. Fit benchmark models (on transformed data)
## Split train test sets
```{r}
Immigrant_ts <- CAN_transform %>% 
  select (Year, Immigrants)

train <- Immigrant_ts  %>% 
    filter (Year <=2018)

test <- Immigrant_ts  %>% 
    filter (Year > 2018)

```

## Fit models
```{r}
# Fit model on train set
models <- train %>%
  model(
    Mean = MEAN(Immigrants),
    Naive = NAIVE(Immigrants),
    Seasonal_Naive = SNAIVE(Immigrants),
    Drift = RW(Immigrants ~ drift())
  )

# Forecast on test set
forecasts <- models %>%
  forecast(test)  

accuracy(models)


# Plot the forecast
forecasts %>%
  autoplot(train, level = NULL) +
  autolayer(Immigrant_ts, Immigrants, colour = "blue") +
  labs(title = "Forecasts vs Actual Data", y = "Immigrants")


```
```{r}
predicted_values <- bench_forecasts$.mean
actual_values <- CAN_transform_test$Immigrants[1:5]

mae <- mean(abs(predicted_values - actual_values))
rmse <- sqrt(mean((predicted_values - actual_values)^2))
mape <- mean(abs((predicted_values - actual_values) / actual_values) * 100)

print(paste("MAE:", mae))
print(paste("RMSE:", rmse))
print(paste("MAPE:", mape))

```



```{r}
# Fit model using Cross-Validation
cv_bench <- CAN_transform_cv %>%
  model(
    Mean = MEAN(Immigrants),
    Naive = NAIVE(Immigrants),
    Seasonal_Naive = SNAIVE(Immigrants),
    Drift = RW(Immigrants ~ drift())
  )


```

# 6. Fit other TS models (NOTE: data has already be transformed!!!)
## Splitting data into train test set

```{r}
# Generate train and test data
CAN_transform_train <- CAN_transform %>% filter(Year <= 2018)
CAN_transform_test <- CAN_transform %>% filter(Year > 2018)
CAN_transform_test

# Plot train and test data
CAN_transform_train  %>%
  autoplot(Immigrants) +
  autolayer(CAN_transform_test, Immigrants, colour = "red") +
  labs(title = "Train and Test Data", y = "adjusted_Value")
```

Set up time series cross-validation 
```{r cache = TRUE}
CAN_transform_cv <- CAN_transform |>
  stretch_tsibble(.init = 30, .step = 1) |>
  filter(.id != max(.id))

CAN_transform_cv 
```

## (a) ARIMA
```{r}
# Fit ARIMA with external regressors
arima_model <- CAN_transform_train %>%
  model(
    ARIMA_Regressors = ARIMA(Immigrants ~  Deaths + Emigrants + ReEmigrants)
  )

# Summary of the fitted model
report(arima_model)
```
```{r}
# Forecasting for the test set period (2019-2023)
arima_forecast <- arima_model %>%
  forecast(new_data = CAN_transform_test)

# Plot the forecast
arima_forecast %>%
  autoplot(CAN_transform, level = NULL) +
  autolayer(CAN_transform_test, Immigrants, colour = "red") +
  labs(
    title = "ARIMA with Regressors Forecast",
    y = "Immigrants",
    x = "Year"
  ) +
  theme_minimal()

# Evaluate forecast accuracy
accuracy_arima <- accuracy(arima_forecast, CAN_transform_test)
accuracy_arima

# Residual diagnostics
arima_model %>%
  gg_tsresiduals()

# Ljung-Box test for residual autocorrelation
augment(arima_model) %>%
  features(.innov, ljung_box, lag = 10)

```
## (b) NNAR
```{r}
# Fit NNAR model with external regressors
nnar_model <- CAN_transform_train %>%
  model(
    NNAR_Regressors = NNETAR(Immigrants ~ Deaths + Emigrants + ReEmigrants, p=1)
  )

# Summary of the fitted NNAR model
report(nnar_model)
```
```{r cache = TRUE}
# Forecast on the test set with external regressors
nnar_forecasts <- nnar_model %>%
  forecast(new_data = CAN_transform_test)

# View the forecasted values
nnar_forecasts

# Plot the forecast
nnar_forecasts %>%
  autoplot(CAN_transform, level = NULL) +
  autolayer(CAN_transform_test, Immigrants, colour = "red") +
  labs(
    title = "NNAR with Regressors Forecast",
    y = "Immigrants",
    x = "Year"
  ) +
  theme_minimal()

# Accuracy of the forecast
nnar_accuracy <- accuracy(nnar_forecasts, CAN_transform_test)

# Print the accuracy results
print(nnar_accuracy)

# Residual diagnostics
nnar_model %>%
  gg_tsresiduals()

# Ljung-Box test for residual autocorrelation
augment(nnar_model) %>%
  features(.innov, ljung_box, lag = 10)
```
## (c) TSLM
```{r}
# Fit TSLM model with external regressors
tslm_model <- CAN_transform_train %>%
  model(
    TSLM_Regressors = TSLM(Immigrants ~ Deaths + Emigrants + ReEmigrants)
  )

# Summary of the fitted TSLM model
report(tslm_model)
```
```{r}
# Forecast on the test set with external regressors
tslm_forecasts <- tslm_model %>%
  forecast(new_data = CAN_transform_test)

# View the forecasted values
tslm_forecasts

# Plot the forecast
tslm_forecasts %>%
  autoplot(CAN_transform, level = NULL) +
  autolayer(CAN_transform_test, Immigrants, colour = "red") +
  labs(
    title = "TSLM with Regressors Forecast",
    y = "Immigrants",
    x = "Year"
  ) +
  theme_minimal()

# Accuracy of the forecast
tslm_accuracy <- accuracy(tslm_forecasts, CAN_transform_test)

# Print the accuracy results
print(tslm_accuracy)

# Residual diagnostics
tslm_model %>%
  gg_tsresiduals()

# Ljung-Box test for residual autocorrelation
augment(tslm_model) %>%
  features(.innov, ljung_box, lag = 10)
```
## (d) PROPHET
```{r cache = TRUE}
# Prepare the data for Prophet
prophet_data <- CAN_transform_train %>%
  select(Year, Immigrants, Deaths, Emigrants, ReEmigrants) %>%
  rename(ds = Year, y = Immigrants)  # Prophet requires 'ds' for time and 'y' for the target variable

# Create Prophet model with external regressors
prophet_model <- prophet(
  yearly.seasonality = TRUE,  # Enable yearly seasonality (you can change this based on your data)
  daily.seasonality = FALSE,
  weekly.seasonality = FALSE
)

# Add external regressors to the model before fitting
prophet_model <- add_regressor(prophet_model, 'Deaths')
prophet_model <- add_regressor(prophet_model, 'Emigrants')
prophet_model <- add_regressor(prophet_model, 'ReEmigrants')

# Fit the model
prophet_model <- fit.prophet(prophet_model, prophet_data)
```
```{r}
# Prepare the test data for forecasting
future_data <- CAN_transform_test %>%
  select(Year, Deaths, Emigrants, ReEmigrants) %>%
  rename(ds = Year)  

# Forecast for the test data (future values)
forecast <- predict(prophet_model, future_data)

# View the forecast results
head(forecast)

# Plot the forecast
plot(prophet_model, forecast)

# Show the model components (trend, seasonality, and the effect of external regressors)
prophet_components <- prophet_plot_components(prophet_model, forecast)
prophet_components
```
```{r cache = TRUE}
actuals <- CAN_transform_test$Immigrants

# Extract the predicted values
predictions <- forecast$yhat

# Calculate MAE
mae <- mean(abs(predictions - actuals))

# Calculate MSE
mse <- mean((predictions - actuals)^2)

# Calculate RMSE
rmse <- sqrt(mse)

# Print the accuracy metrics
cat("MAE:", mae, "\n")
cat("MSE:", mse, "\n")
cat("RMSE:", rmse, "\n")
```


















